---
title: "Cleaning"
output: html_document
date: "2023-08-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Up Packages

```{r}
rm(list = ls())

library(ggplot2)
library(ggspatial)
library(tidyr)
library(dplyr)
library(sf)
library(raster)
library(nngeo)
library(geobr)
library(terra)
```

# Loading Up Data:

```{r}
mun_census_1872 <- read_municipality(code_muni = "all",year = 1872, simplified = TRUE, showProgress = TRUE)

setwd("~/OneDrive - University of Illinois - Urbana/Research Data/Brazil/Census 1872")
census_1872 <- read_xlsx("municipalities_1872_census.xlsx")
census_names_1872 <- read_xlsx("CensusToShapeNames.xlsx")

sugarcane <- 
  terra::rast("~/OneDrive - University of Illinois - Urbana/Research Data/FAO_GAEZ/sugarcane_lowinput_CO2.tiff")
```

# Cleaning Data:

```{r}
census_1872_clean <- merge(census_1872, census_names_1872, by = c("MunicId", "ProvId"))
```

```{r}
census_1872_clean$abbrev_state <- ""
census_1872_clean$abbrev_state <- 
  ifelse(census_1872_clean$PrimeiroDeProvincia == "Alagoas","AL", census_1872_clean$abbrev_state)
census_1872_clean$abbrev_state <- 
  ifelse(census_1872_clean$PrimeiroDeProvincia == "Amazonas","AM", census_1872_clean$abbrev_state)
census_1872_clean$abbrev_state <- 
  ifelse(census_1872_clean$PrimeiroDeProvincia == "Bahia","BA", census_1872_clean$abbrev_state)
census_1872_clean$abbrev_state <- 
  ifelse(census_1872_clean$PrimeiroDeProvincia == "Ceará", "CE", census_1872_clean$abbrev_state)
census_1872_clean$abbrev_state <- 
  ifelse(census_1872_clean$PrimeiroDeProvincia == "Espírito Santo","ES", census_1872_clean$abbrev_state)
census_1872_clean$abbrev_state <- 
  ifelse(census_1872_clean$PrimeiroDeProvincia == "Goyaz","GO", census_1872_clean$abbrev_state)
census_1872_clean$abbrev_state <- 
  ifelse(census_1872_clean$PrimeiroDeProvincia == "Maranhão","MA", census_1872_clean$abbrev_state)
census_1872_clean$abbrev_state <- 
  ifelse(census_1872_clean$PrimeiroDeProvincia == "Minas Geraes","MG", census_1872_clean$abbrev_state)
census_1872_clean$abbrev_state <- 
  ifelse(census_1872_clean$PrimeiroDeProvincia == "Matto Grosso","MT", census_1872_clean$abbrev_state)
census_1872_clean$abbrev_state <- 
  ifelse(census_1872_clean$PrimeiroDeProvincia == "Pará","PA", census_1872_clean$abbrev_state)
census_1872_clean$abbrev_state <- 
  ifelse(census_1872_clean$PrimeiroDeProvincia == "Parahyba","PB", census_1872_clean$abbrev_state)
census_1872_clean$abbrev_state <- 
  ifelse(census_1872_clean$PrimeiroDeProvincia == "Pernambuco","PE", census_1872_clean$abbrev_state)
census_1872_clean$abbrev_state <- 
  ifelse(census_1872_clean$PrimeiroDeProvincia == "Piauhy","PI", census_1872_clean$abbrev_state)
census_1872_clean$abbrev_state <- 
  ifelse(census_1872_clean$PrimeiroDeProvincia == "Paraná","PR", census_1872_clean$abbrev_state)
census_1872_clean$abbrev_state <- 
  ifelse(census_1872_clean$PrimeiroDeProvincia == "Rio de Janeiro","RJ", census_1872_clean$abbrev_state)
census_1872_clean$abbrev_state <- 
  ifelse(census_1872_clean$PrimeiroDeProvincia == "Rio Grande do Norte","RN", census_1872_clean$abbrev_state)
census_1872_clean$abbrev_state <- 
  ifelse(census_1872_clean$PrimeiroDeProvincia == "Rio Grande do Sul","RS", census_1872_clean$abbrev_state)
census_1872_clean$abbrev_state <- 
  ifelse(census_1872_clean$PrimeiroDeProvincia == "Santa Catharina","SC", census_1872_clean$abbrev_state)
census_1872_clean$abbrev_state <- 
  ifelse(census_1872_clean$PrimeiroDeProvincia == "Sergipe","SE", census_1872_clean$abbrev_state)
census_1872_clean$abbrev_state <- 
  ifelse(census_1872_clean$PrimeiroDeProvincia == "São Paulo","SP", census_1872_clean$abbrev_state)
census_1872_clean$abbrev_state <- 
  ifelse(census_1872_clean$PrimeiroDeProvincia == "Municipio Neutro ( Município da Corte)","RJ",
         census_1872_clean$abbrev_state)

### 
census_1872_clean$PrimeiroDeProvincia <- 
  ifelse(census_1872_clean$PrimeiroDeProvincia == "Municipio Neutro ( Município da Corte)",
         "Municipio Neutro",
         census_1872_clean$PrimeiroDeProvincia)
```

```{r}
clean_1872 <- merge(mun_census_1872, census_1872_clean, by = c("name_muni", "abbrev_state"))
clean_1872$sugarcane <- exact_extract(sugarcane, clean_1872, 'mean', progress = FALSE)
```

```{r}
clean_1872$free_slave <- 
  100*clean_1872$SomaDeSOMA_E_Almas/(clean_1872$SomaDeSOMA_L_Almas + clean_1872$SomaDeSOMA_E_Almas)

clean_1872$literacy_rate <- 
  100*(clean_1872$`SomaDeSOMA_L_Sabem Ler e Escrever`/
         (clean_1872$SomaDeH_LIVRES_Almas + clean_1872$SomaDeM_LIVRES_Almas))

clean_1872$s_literacy_rate <-
  100*(clean_1872$`SomaDeSOMA_E_Sabem Ler e Escrever`/clean_1872$SomaDeSOMA_E_Almas)

clean_1872$farmers <-
  100*(clean_1872$SomaDeSOMA_G_Lavradores + clean_1872$SomaDeSOMA_G_Criadores)/(clean_1872$SomaDeSOMA_L_Almas + clean_1872$SomaDeSOMA_E_Almas)

clean_1872$farmers_slaves <-
  100*(clean_1872$SomaDeSOMA_E_Lavradores + clean_1872$SomaDeSOMA_E_Criadores)/(clean_1872$SomaDeSOMA_E_Almas)

clean_1872$industrial_workers <-
  100*(clean_1872$`SomaDeSOMA_G_Manufatureiros e fabricantes` + clean_1872$`SomaDeSOMA_G_Comerciantes, guarda-livros e caixeiros`)/(clean_1872$SomaDeSOMA_L_Almas + clean_1872$SomaDeSOMA_E_Almas)

clean_1872$gender_disparity <-
  (clean_1872$SomaDeH_LIVRES_Almas/clean_1872$SomaDeM_LIVRES_Almas)

clean_1872$slave_gender_disparity <-
  (clean_1872$SomaDeH_ESCR_Almas/clean_1872$SomaDeM_ESCR_Almas)

clean_1872$total_slaves <- clean_1872$SomaDeSOMA_E_Almas

#clean_1872$p_agriculture <-
#  100*(parishes_1872$SOMA_L_Agriculture/parishes_1872$SOMA_L_Almas)
```

```{r}
ggplot() +
  geom_sf(data = clean_1872, aes(fill = free_slave)) +
  scale_fill_binned(type = "viridis", n.breaks = 10) + 
  theme_bw()

ggplot() +
  geom_sf(data = clean_1872, aes(fill = sugarcane)) +
  scale_fill_binned(type = "viridis", n.breaks = 10) + 
  theme_bw()
```

```{r}
sesmarias_pts <- read_xlsx("/Users/vinicius/Library/CloudStorage/OneDrive-UniversityofIllinois-Urbana/Research/Projects/Sesmarias Brazil/Data/sesmarias_georeferenced.xlsx")

mun_1872 <- read_municipality(code_muni = "all",year = 1872, simplified = TRUE, showProgress = TRUE)
mun_2010 <- read_municipality(code_muni = "all",year = 2010, simplified = TRUE, showProgress = TRUE)

mun_2010$sugarcane <- exact_extract(sugarcane, mun_2010, 'mean', progress = FALSE)

sesmarias_pts$Latitude <- as.numeric(gsub('°','',sesmarias_pts$Latitude))
sesmarias_pts$Longitude <- as.numeric(gsub('°','',sesmarias_pts$Longitude))

# brazil <- 

sesmarias_pts <- subset(sesmarias_pts, 
                        is.na(sesmarias_pts$Latitude) == F & is.na(sesmarias_pts$Longitude) == F)
sesmarias_pts <- st_as_sf(sesmarias_pts, coords = c("Longitude", "Latitude"), 
                          crs = st_crs(mun_1872))
```

# Merge that to the database to get the sizes and other stuff:

-   Manually check the economic activities

```{r}
p <- st_intersects(clean_1872, sesmarias_pts)
p[lengths(p) > 0] <- 1
p[lengths(p) == 0] <- 0

clean_1872$sesmarias_presence <- unlist(p)

p <- st_intersects(clean_1872, sesmarias_pts)
p[lengths(p) > 0] <- lengths(p)
p[lengths(p) == 0] <- 0

clean_1872$sesmarias_intensity <- unlist(p)

# 2010:

mun_2010_NE <- subset(mun_2010, abbrev_state %in% c("PB", "PE", "RN", "AL"))

p <- st_intersects(mun_2010_NE, sesmarias_pts)
p[lengths(p) > 0] <- 1
p[lengths(p) == 0] <- 0

mun_2010_NE$sesmarias_presence <- unlist(p)
```

```{r}
ggplot() +
  geom_sf(data = mun_1872, fill = "white") +
  geom_sf(data = subset(clean_1872,  abbrev_state %in% c("PB", "PE", "RN", "AL", "CE")), 
          aes(fill = free_slave)) +
  scale_fill_binned(type = "viridis", n.breaks = 10) + 
  geom_sf(data = sesmarias_pts, size = 0.25, color = "red") +
  coord_sf(ylim = c(-2,-15), xlim = c(-34,-45), expand = FALSE) + 
  labs(fill = "Proportion of Slaves \n to Total Population (%)") +
  theme_bw() +
  theme(#legend.position = "bottom", 
        plot.title = element_text(hjust = 0.5))
ggsave("/Users/vinicius/Library/CloudStorage/OneDrive-UniversityofIllinois-Urbana/Research/Projects/Sesmarias Brazil/Figures/01. Maps/slave_proportion.png")

ggplot() +
  geom_sf(data = mun_1872, fill = "white") +
  geom_sf(data = subset(clean_1872,  abbrev_state %in% c("PB", "PE", "RN", "AL", "CE")), 
          aes(fill = sugarcane)) +
  scale_fill_binned(type = "viridis", n.breaks = 10) + 
  geom_sf(data = sesmarias_pts, size = 0.25, color = "red") +
  coord_sf(ylim = c(-2,-15), xlim = c(-34,-45), expand = FALSE) + 
  labs(fill = "Maximum Sugar \n Cane Production") +
  theme_bw() +
  theme(# legend.position = "bottom", 
        plot.title = element_text(hjust = 0.5))
ggsave("/Users/vinicius/Library/CloudStorage/OneDrive-UniversityofIllinois-Urbana/Research/Projects/Sesmarias Brazil/Figures/01. Maps/sugarcane_production.png")

ggplot() +
  geom_sf(data = mun_1872, fill = "white") +
  geom_sf(data = subset(clean_1872,  abbrev_state %in% c("PB", "PE", "RN", "AL", "CE")), 
          aes(fill = farmers)) +
  scale_fill_binned(type = "viridis", n.breaks = 10) + 
  geom_sf(data = sesmarias_pts, size = 0.25, color = "red") +
  coord_sf(ylim = c(-2,-15), xlim = c(-34,-45), expand = FALSE) + 
  labs(fill = "Proportion of \n Farmers to Pop.") +
  theme_bw() +
  theme(# legend.position = "bottom", 
        plot.title = element_text(hjust = 0.5))
ggsave("/Users/vinicius/Library/CloudStorage/OneDrive-UniversityofIllinois-Urbana/Research/Projects/Sesmarias Brazil/Figures/01. Maps/farmers_share.png")

ggplot() +
  geom_sf(data = mun_1872, fill = "white") +
  geom_sf(data = subset(clean_1872,  abbrev_state %in% c("PB", "PE", "RN", "AL", "CE")), 
          aes(fill = farmers_slaves)) +
  scale_fill_binned(type = "viridis", n.breaks = 10) + 
  geom_sf(data = sesmarias_pts, size = 0.25, color = "red") +
  coord_sf(ylim = c(-2,-15), xlim = c(-34,-45), expand = FALSE) + 
  labs(fill = "Proportion of Slave \n Farmers to Slave Pop.") +
  theme_bw() +
  theme(# legend.position = "bottom", 
        plot.title = element_text(hjust = 0.5))
ggsave("/Users/vinicius/Library/CloudStorage/OneDrive-UniversityofIllinois-Urbana/Research/Projects/Sesmarias Brazil/Figures/01. Maps/slave_farmers_share.png")

ggplot() +
  geom_sf(data = mun_1872, fill = "white") +
  geom_sf(data = subset(clean_1872,  abbrev_state %in% c("PB", "PE", "RN", "AL", "CE")), 
          aes(fill = industrial_workers)) +
  scale_fill_binned(type = "viridis", n.breaks = 10) + 
  geom_sf(data = sesmarias_pts, size = 0.25, color = "red") +
  coord_sf(ylim = c(-2,-15), xlim = c(-34,-45), expand = FALSE) + 
  labs(fill = "Proportion of Industry \n Workers Pop.") +
  theme_bw() +
  theme(# legend.position = "bottom", 
        plot.title = element_text(hjust = 0.5))
ggsave("/Users/vinicius/Library/CloudStorage/OneDrive-UniversityofIllinois-Urbana/Research/Projects/Sesmarias Brazil/Figures/01. Maps/industry_share.png")

ggplot() +
  geom_sf(data = mun_1872, fill = "white") +
  geom_sf(data = subset(clean_1872,  abbrev_state %in% c("PB", "PE", "RN", "AL", "CE")), 
          aes(fill = gender_disparity)) +
  scale_fill_binned(type = "viridis", n.breaks = 10) + 
  geom_sf(data = sesmarias_pts, size = 0.25, color = "red") +
  coord_sf(ylim = c(-2,-15), xlim = c(-34,-45), expand = FALSE) + 
  labs(fill = "Ratio of Men \n to Women.") +
  theme_bw() +
  theme(# legend.position = "bottom", 
        plot.title = element_text(hjust = 0.5))
ggsave("/Users/vinicius/Library/CloudStorage/OneDrive-UniversityofIllinois-Urbana/Research/Projects/Sesmarias Brazil/Figures/01. Maps/gender_ratio.png")

ggplot() +
  geom_sf(data = mun_1872, fill = "white") +
  geom_sf(data = subset(clean_1872,  abbrev_state %in% c("PB", "PE", "RN", "AL", "CE")), 
          aes(fill = slave_gender_disparity)) +
  scale_fill_binned(type = "viridis", n.breaks = 10) + 
  geom_sf(data = sesmarias_pts, size = 0.25, color = "red") +
  coord_sf(ylim = c(-2,-15), xlim = c(-34,-45), expand = FALSE) + 
  labs(fill = "Ratio of Slave Men \n to Women.") +
  theme_bw() +
  theme(# legend.position = "bottom", 
        plot.title = element_text(hjust = 0.5))
ggsave("/Users/vinicius/Library/CloudStorage/OneDrive-UniversityofIllinois-Urbana/Research/Projects/Sesmarias Brazil/Figures/01. Maps/gender_ratio_slaves.png")

ggplot() +
  geom_sf(data = mun_1872, fill = "white") +
  geom_sf(data = subset(clean_1872,  abbrev_state %in% c("PB", "PE", "RN", "AL", "CE")), 
          aes(fill = log(total_slaves))) +
  scale_fill_binned(type = "viridis", n.breaks = 10) + 
  geom_sf(data = sesmarias_pts, size = 0.25, color = "red") +
  coord_sf(ylim = c(-2,-15), xlim = c(-34,-45), expand = FALSE) + 
  labs(fill = "Log(Total Slaves)") +
  theme_bw() +
  theme(# legend.position = "bottom", 
        plot.title = element_text(hjust = 0.5))
ggsave("/Users/vinicius/Library/CloudStorage/OneDrive-UniversityofIllinois-Urbana/Research/Projects/Sesmarias Brazil/Figures/01. Maps/total_slaves.png")
```

```{r}
ggplot(data = subset(clean_1872,  abbrev_state %in% c("PB", "PE", "RN", "AL", "CE")), 
       aes(x = sugarcane, y = free_slave)) +
  geom_point(color = "black", fill = "red") +
  geom_smooth(method='lm') + 
  labs(y = "Percentage of Slaves to Total Population (%)", x = "Potential Sugarcane Production",
       title = "Relationship between Slavery and Sugarcane in Northeastern States - Brazil") +
  theme_bw() +
  theme(legend.position = "bottom", 
        plot.title = element_text(hjust = 0.5))
ggsave("/Users/vinicius/Library/CloudStorage/OneDrive-UniversityofIllinois-Urbana/Research/Projects/Sesmarias Brazil/Figures/02. Graph/sugarcane_slavery.png")

ggplot(data = subset(clean_1872,  abbrev_state %in% c("PB", "PE", "RN", "AL", "CE")), 
       aes(x = sugarcane, y = literacy_rate)) +
  geom_point(color = "black", fill = "red") +
  geom_smooth(method='lm') + 
  labs(y = "Literacy Rate (%)", x = "Potential Sugarcane Production",
       title = "Relationship between Sugarcane and Literacy in Northeastern States - Brazil") +
  theme_bw() +
  theme(legend.position = "bottom", 
        plot.title = element_text(hjust = 0.5))
ggsave("/Users/vinicius/Library/CloudStorage/OneDrive-UniversityofIllinois-Urbana/Research/Projects/Sesmarias Brazil/Figures/02. Graph/sugarcane_literacy.png")

ggplot(data = subset(clean_1872,  abbrev_state %in% c("PB", "PE", "RN", "AL", "CE")), 
       aes(x = free_slave, y = literacy_rate)) +
  geom_point(color = "black", fill = "red") +
  geom_smooth(method='lm') + 
  labs(y = "Literacy Rate (%)", x = "Percentage of Slaves to Total Population (%)",
       title = "Relationship between Slavery and Literacy in Northeastern States - Brazil") +
  theme_bw() +
  theme(legend.position = "bottom", 
        plot.title = element_text(hjust = 0.5))
ggsave("/Users/vinicius/Library/CloudStorage/OneDrive-UniversityofIllinois-Urbana/Research/Projects/Sesmarias Brazil/Figures/02. Graph/slavery_literacy.png")
```

```{r}
ggplot() +
  geom_sf(data = mun_2010, fill = "white") +
  geom_sf(data = subset(mun_2010, abbrev_state %in% c("PB", "PE", "RN", "AL", "CE")), fill = "lightblue") +
  geom_sf(data = sesmarias_pts, size = 0.25, color = "red") +
  coord_sf(ylim = c(-2,-15), xlim = c(-34,-45), expand = FALSE) + 
  theme_bw()
```

# Summary Tables:

```{r}

```

# Regressions:

```{r}
summary(lm_robust(sesmarias_presence ~ log(sugarcane), 
                  fixed_effects = ~abbrev_state, 
                  data = subset(clean_1872,  abbrev_state %in% c("PB", "PE", "RN", "AL", "CE"))))

summary(lm_robust(free_slave ~ sesmarias_intensity, 
                  fixed_effects = ~abbrev_state, 
                  data = subset(clean_1872,  abbrev_state %in% c("PB", "PE", "RN", "AL", "CE"))))
```

# 2010 Data:

```{r}
race2010 <- read_xlsx("~/OneDrive - University of Illinois - Urbana/Research Data/Brazil/Census 2010/race2010census_clean.xlsx")

gini_2010 <- 
  read.csv("~/OneDrive - University of Illinois - Urbana/Research Data/Brazil/DataSUS/Dados/GINI/ginibr_clean.csv")

gdp_2010 <- read_xlsx("~/OneDrive - University of Illinois - Urbana/Research/Projects/Amazon/gdp/GDP1999-2017.xlsx")

mun_2010 <- merge(race2010, mun_2010, by = c("code_muni"))
```

```{r}
ggplot() +
  geom_sf(data = mun_1872, fill = "white") +
  geom_sf(data = subset(clean_1872,  abbrev_state %in% c("PB", "PE", "RN", "AL", "CE")), 
          aes(fill = free_slave)) +
  scale_fill_binned(type = "viridis", n.breaks = 10) + 
  geom_sf(data = sesmarias_pts, size = 0.25, color = "red") +
  coord_sf(ylim = c(-2,-15), xlim = c(-34,-45), expand = FALSE) + 
  labs(fill = "Proportion of Slaves \n to Total Population (%)") +
  theme_bw()

ggplot() +
  geom_sf(data = mun_2010$geom, fill = "white") +
  geom_sf(data = subset(mun_2010,  abbrev_state %in% c("PB", "PE", "RN", "AL", "CE")), 
          aes(fill = Preta)) +
  scale_fill_binned(type = "viridis", n.breaks = 10) + 
  geom_sf(data = sesmarias_pts, size = 0.25, color = "red") +
  coord_sf(ylim = c(-2,-15), xlim = c(-34,-45), expand = FALSE) + 
  theme_bw()
```

# Testing Balance

Have to add observables, instead of outcomes here. Stuff like potential for sugarcane, slope, elevation, distance to river, distance to the coast, longitude, and latitude.

```{r}
summary_1872 <- subset(clean_1872,  abbrev_state %in% c("PB", "PE", "RN", "AL"))

summary_1872 <- summary_1872[c("sesmarias_presence",
                               "sugarcane", 
                               "literacy_rate", 
                               "SomaDeSOMA_G_Almas",
                             "free_slave", 
                             "industrial_workers", 
                             "farmers_slaves",
                             "farmers")]

summary_1872$within <- 
  ifelse(summary_1872$sesmarias_presence > 0, 
         "At least 1 Land Grant", 
         "No Land Grants")

summary_1872 <- dplyr::rename(summary_1872, 
                              "Land Grants Presence" = sesmarias_presence,
                               "Potential Sugarcane Output (FAO)" = sugarcane, 
                               "Literacy Rate (%)" = literacy_rate, 
                               "Total Population" = SomaDeSOMA_G_Almas,
                               "Slave Population as Total (%)" = free_slave, 
                               "Proportion of Industrial Workers (%)" = industrial_workers, 
                              "Proportion of Slaves working in Farming (%)" = farmers_slaves,
                              "Proportion of Farmers (%)" = farmers)




datasummary_balance(~within, summary_1872)
```

```{r}
summary(lm_robust(sesmarias_presence ~ log(sugarcane+1),
                  fixed_effects = ~abbrev_state,
                  data = mun_2010_NE))
```
